---
title: "STAT433_HW1"
subtitle: "Olivia Plamann"
output: github_document
---

```{r, message=FALSE}
library(dplyr)
library(nycflights13)
library(ggplot2)
```

#### Question 1:
##### 1a). How many flights have a missing dep_time? 
```{r}
#1a).
missing_dep_time <- flights %>%
  filter(is.na(dep_time))

cat(length(missing_dep_time$dep_time), "flights have a missing dep_time")
```

##### 1b). What other variables are missing? 
The columns dep_time, dep_delay, arr_time, and arr_delay are also all missing.  Tailnum has some missing values.
```{r}
#1b).
missing_dep_time %>%
  summarize(num_missing_dep_time = sum(is.na(dep_time)),
            num_missing_dep_delay = sum(is.na(dep_delay)),
            num_missing_arr_time = sum(is.na(arr_time)),
            num_missing_arr_delay = sum(is.na(arr_delay)),
            num_missing_tailnum = sum(is.na(tailnum)))
```

##### 1c). What might these rows represent?
These rows might represent cancelled flights or flights that never actually took off.

#### Question 2: Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.
```{r}
#2).
minutes_since_midnight <- flights %>%
  select(dep_time, sched_dep_time) %>%
  mutate(hour_dep = dep_time %/% 100,
         minutes_dep = dep_time - (hour_dep * 100),
         minutes_since_dep = ((hour_dep*60)+minutes_dep) %% 1440,
         hour_sched = sched_dep_time %/% 100,
         minutes_sched = sched_dep_time - (hour_sched * 100),
         minutes_since_sched = ((hour_sched*60)+minutes_sched) %% 1440) %>%
  select(dep_time, minutes_since_dep, sched_dep_time, minutes_since_sched)

minutes_since_midnight

```

#### Question 3: 
##### 3a). Look at the number of cancelled flights per day. Is there a pattern? 
```{r, message=FALSE}
# 3a).
delayed_count <- flights %>%
  mutate(cancelled = (is.na(dep_delay) | is.na(arr_delay))) %>%
  group_by(year, month, day) %>%
  summarize(count_cancel = sum(cancelled),
            count_total = n())

ggplot(delayed_count, aes(x=count_total, y=count_cancel)) +
         geom_point()
```
The pattern is that the number of cancelled flights seems to increase as the total number of flights increases.

##### 3b). Is the proportion of cancelled flights related to the average delay? Use multiple dyplr operations, all on one line, concluding with ggplot(aes(x= ,y=)) + geom_point()

```{r, message=FALSE}
# 3b).
# DOING FOR BOTH ARRIVAL AND DEPARTURE AS IT IS NOT SPECIFIED:
# Departure
flights %>%
  mutate(cancelled = (is.na(dep_delay) | is.na(arr_delay))) %>%
  group_by(year, month, day) %>%
  summarize(prop_cancel = mean(cancelled),
            avg_dep_delay = mean(dep_delay, na.rm=TRUE)) %>%
  ggplot(aes(x=avg_dep_delay,y=prop_cancel)) + geom_point()

# Arrival
flights %>%
  mutate(cancelled = (is.na(dep_delay) | is.na(arr_delay))) %>%
  group_by(year, month, day) %>%
  summarize(prop_cancel = mean(cancelled),
            avg_arr_delay = mean(arr_delay, na.rm=TRUE)) %>%
  ggplot(aes(x=avg_arr_delay,y=prop_cancel)) + geom_point()
```


